---
title: "Replicate MRW in full"
author: '@lachlandeer & @julianlanger'
date: "April 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There probably won't be too much expository text.
Essentially we want to replicate the entirety of the empirical component of Mankiw, Romer and Weil (1992)

## Libraries

```{r, warning=FALSE, message=FALSE}
library(haven)      # reading stata data
library(dplyr)      # data manipulation
library(tibble)     # nicer dataframes
library(stargazer)  # tables
library(ggplot2)    # graphs
library(car)        # linear hypotheses
```


## Load data
```{r}
mrw_data <- read_dta("../src/data/mrw.dta")
```

Structure?

```{r}
str(mrw_data)
skimr::skim(mrw_data)
```

## Meaningful variable names

```{r}

mrw_data <- mrw_data %>% 
              rename(non_oil = n, 
                     oecd = o,
                     intermediate = i,
                     gdp_60 = rgdpw60,
                     gdp_85 = rgdpw85,
                     gdp_growth_60_85 = gdpgrowth,
                     pop_growth_60_85 = popgrowth,
                     inv_gdp = i_y,
                     school = school)

```

## Variable Creation

```{r}
delta_gamma <- 0.05

mrw_data <- mrw_data %>% 
                  mutate(ln_gdp_85 = log(gdp_85),
                         ln_gdp_60 = log(gdp_60),
                         ln_gdp_growth = ln_gdp_85 - ln_gdp_60,
                         ln_inv_gdp = log(inv_gdp/100),
                         non_oil = factor(non_oil),
                         intermediate = factor(intermediate),
                         oecd = factor(oecd),
                         ln_ndg = log(pop_growth_60_85/100 + delta_gamma),
                         ln_school = log(school/100)) %>% 
                  select(country, ln_gdp_85, ln_gdp_60, ln_inv_gdp, 
                         non_oil, intermediate, oecd,
                         ln_ndg, ln_school, gdp_growth_60_85, ln_gdp_growth)

```

## Samples

```{r}
oecd         <- mrw_data %>% 
                    filter(oecd == 1)
intermediate <- mrw_data %>% 
                    filter(intermediate == 1)
non_oil      <- mrw_data %>% 
                    filter(non_oil == 1)
```

## Table 1

### Unrestricted Regression

```{r}
solow_oecd <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = oecd)
solow_int  <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = intermediate)
solow_noil <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = non_oil)
```

#### Regression Table

```{r, results = 'asis'}
stargazer(solow_noil, solow_int, solow_oecd, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)", 
                               "log(n+delta+g)", 
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 1 - Unrestricted Models",
          style = "qje")
```


### Restricted Regression

```{r}
solow_oecd_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = oecd)
solow_int_r  <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = intermediate)
solow_noil_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = non_oil)
```

#### Regression Table

```{r, results = 'asis'}
stargazer(solow_noil_r, solow_int_r, solow_oecd_r, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)- log(n+delta+g)", 
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 1 - Restricted Models",
          style = "qje")
```

### Tables to modelsummary

```{r}
library(modelsummary)
library(kableExtra)
```


```{r}
models <- list(
    solow_oecd <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = oecd),
    solow_int  <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = intermediate),
    solow_noil <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = non_oil),
    solow_oecd_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = oecd),
    solow_int_r  <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = intermediate),
    solow_noil_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = non_oil)
)

names(models) <- c("(1)", "(2)", "(3)", "(4)", "(5)", "(6)")

custom_row <- tibble::tribble(~term, ~"(1)", ~"(2)", ~"(3)", ~"(4)", ~"(5)", ~"(6)",
                              "Restricted Model", "No", "No", "No", "Yes", "Yes", "Yes")

attr(custom_row, 'position') <- c(9)

f_int <- function(x) format(round(x, 2), big.mark=",", nsmall = 0) # integers
f_dec <- function(x) format(round(x, 2), big.mark=",", nsmall = 2) # floats

gm <- list(
  list("raw" = "nobs", "clean" = "N", "fmt" = f_int),
  list("raw" = "r.squared", "clean" = "$R^2$", "fmt" = f_dec))

tab <- modelsummary(models,
             fmt = "%.2f",
             coef_map = c(    "ln_inv_gdp" = "log(I / GDP)",
                              "ln_ndg" = "log(n + g + $\\delta$)",
                              "I(ln_inv_gdp - ln_ndg)"=  "log(I / GDP) - log(n + g + $\\delta$)",
                              '(Intercept)' = 'Constant'
                               ),
             statistic = "std.error",
             stars = FALSE,
             stars_note = FALSE,
             gof_omit = 'DF|Deviance|Adj|AIC|BIC|Log.|F',
             gof_map = gm,
             add_rows = custom_row,
             output = 'latex_tabular',
             escape = FALSE # render latex as latex- needs FALSE
             )
tab 
tab2 <- tab %>%
    add_header_above(c(" " =1, "Non-Oil" =1, "Intermediate" =1 , "OECD"=1,
                            "Non-Oil"=1, "Intermediate"=1, "OECD"=1)) %>%
    add_header_above(c(" " =1, "Log(GDP '85) - Log(GDP '60)" =6)) %>%
    row_spec(9, extra_latex_after = "\\midrule") %>%
    save_kable(file = "table.tex", self_contained = FALSE)

```

### Test of Restriction

```{r}
linearHypothesis(solow_oecd, "ln_inv_gdp = - ln_ndg")
linearHypothesis(solow_noil, "ln_inv_gdp = - ln_ndg")
linearHypothesis(solow_int, "ln_inv_gdp = - ln_ndg")
```

### Implied alpha 


```{r}
alpha_solow_noil <- round(coef(solow_noil_r)[2] / (1 + coef(solow_noil_r)[2]), digits = 2)
print(paste("Implied Alpha (Non oil):", alpha_solow_noil))

alpha_solow_int <- round(coef(solow_int_r)[2] / (1 + coef(solow_int_r)[2]), digits = 2)
print(paste("Implied Alpha (Intermediate):", alpha_solow_int))

alpha_solow_oecd <- round(coef(solow_oecd_r)[2] / (1 + coef(solow_oecd_r)[2]), digits = 2)
print(paste("Implied Alpha (OECD):", alpha_solow_oecd))
```

## Table 2

### Unrestricted Model
```{r}
augsolow_oecd <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = oecd)
augsolow_int  <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = intermediate)
augsolow_noil <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = non_oil)
```

#### Regression Table
```{r, results = 'asis'}
stargazer(augsolow_noil, augsolow_int, augsolow_oecd, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)", 
                               "log(n+delta+g)",
                               "log(school)",
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 2 - Unrestricted Models",
          style = "qje")
```

#### Restricted Regression

```{r}
augsolow_oecd_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = oecd)
augsolow_int_r  <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = intermediate)
augsolow_noil_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = non_oil)
```

### Regression Table
```{r, results = 'asis'}
stargazer(augsolow_noil_r, augsolow_int_r, augsolow_oecd_r, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)- log(n+delta+g)", 
                               "log(school)- log(n+delta+g)",
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 2 - Restricted Models",
          style = "qje")
```

### Test of Restriction
```{r}
linearHypothesis(augsolow_oecd, "ln_inv_gdp + ln_ndg + ln_school = 0")
linearHypothesis(augsolow_noil, "ln_inv_gdp + ln_ndg + ln_school = 0")
linearHypothesis(augsolow_int, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

## Table 3

```{r}
ucc_noil <- lm(ln_gdp_growth ~ ln_gdp_60, data = non_oil)
ucc_int  <- lm(ln_gdp_growth ~ ln_gdp_60, data = intermediate)
ucc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60, data = oecd)
```

### Regression Table
```{r, results = 'asis'}
stargazer(ucc_noil, ucc_int, ucc_oecd, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                              "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 3 - Unconditional Convergence",
          style = "qje")
```

## Table 4
```{r}
cc_noil <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = non_oil)
cc_int  <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = intermediate)
cc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = oecd)
```

### Regression Table
```{r, results = 'asis'}
stargazer(cc_noil, cc_int, cc_oecd, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP)", 
                               "log(n+delta+g)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 4 - Conditional Convergence",
          style = "qje")
```

## Table 5

```{r}
augcc_noil <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = non_oil)
augcc_int  <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = intermediate)
augcc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = oecd)
```

### Regression Table
```{r, results = 'asis'}
stargazer(augcc_noil, augcc_int, augcc_oecd, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP)", 
                               "log(n+delta+g)",
                               "log(school)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 5 - Augmented Conditional Convergence",
          style = "qje")
```

## Table 6
```{r}
augcc_noil_r <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = non_oil)
augcc_int_r  <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = intermediate)
augcc_oecd_r <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = oecd)
```

### Regression Table
```{r, results = 'asis'}
stargazer(augcc_noil_r, augcc_int_r, augcc_oecd_r, type = "html",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP) - log(n+delta+g)", 
                               "log(school) - log(n+delta+g)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 6 - Restricted Augmented Conditional Convergence",
          style = "qje")
```

### Test of Restriction
```{r}
linearHypothesis(augcc_oecd, "ln_inv_gdp + ln_ndg + ln_school = 0")
linearHypothesis(augcc_noil, "ln_inv_gdp + ln_ndg + ln_school = 0")
linearHypothesis(augcc_int, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

## Figure 1

Notice that the actual axis labels in the paper are a little misleading. 
These actually replicate the figure.

### Top Panel

```{r}
ggplot(intermediate, aes(x = ln_gdp_60, y = ln_gdp_growth)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("A: Unconditional") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```

### Middle Panel

Takes a little more work, we need the residuals 

```{r}
y2 <- lm(ln_gdp_growth ~  ln_inv_gdp + ln_ndg, data = intermediate)$residuals
x2 <- lm(ln_gdp_60 ~  ln_inv_gdp + ln_ndg, data = intermediate)$residuals

panel_b <- tibble(y2, x2)
```

Now the plot:

```{r}
ggplot(panel_b, aes(x = x2, y = y2)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("B: Conditional on Saving and Population Growth") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```

### Bottom Panel
Again need the residuals 

```{r}
y3 <- lm(ln_gdp_growth ~  ln_inv_gdp + ln_ndg + ln_school, data = intermediate)$residuals
x3 <- lm(ln_gdp_60 ~  ln_inv_gdp + ln_ndg + ln_school, data = intermediate)$residuals

panel_c <- tibble(y3, x3)
```

Now the plot:

```{r}
ggplot(panel_c, aes(x = x3, y = y3)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("C: Conditional on Saving and Population Growth and Human Capital") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```